---
title: "Figures"
format: html
---

The code in this script generates figures for the manuscript.

::: callout-note
### Source code

The full R script is available at:

-   [`R/figures.R`](https://github.com/VagishHemmige/Cryptococcus-cost_analysis/blob/master/R/Figures.R)

This R script file is itself reliant on the following helper files:

-   [`R/setup.R`](https://github.com/VagishHemmige/Cryptococcus-cost_analysis/blob/master/R/setup.R)
-   [`R/functions.R`](https://github.com/VagishHemmige/Cryptococcus-cost_analysis/blob/master/R/functions.R)
:::

## Final flowchart

The following code uses the `flowchart` package to create a STROBE flowchart for the study.

```{r, eval=FALSE}


#Final STROBE flowchart
patients_fc_matched%>%
  fc_draw()%>%
  fc_export("figures/flowchart.svg",
            width=20,
            height=10)



```

## Beeswarm plot of total inflated costs

The following code uses the `ggbeeswarm` add-on to `ggplot2` to compare cases and controls in terms of total inflated costs post-index.  No attempt is made to adjust for censoring or death.

```{r, eval=FALSE}

#GGplot that is beeswarm of adjusted total costs
cost_inflated%>%
  mutate(IN_CLM_365d_cost_adjusted_total=ifelse(IN_CLM_365d_cost_adjusted_total==0, 1, IN_CLM_365d_cost_adjusted_total))%>%
  ggplot()+
  geom_quasirandom(mapping=aes(x=patient_type, y=IN_CLM_365d_cost_adjusted_total), alpha=0.1)+
  scale_y_log10()+
  theme_classic()


```

## Month-by-month plots

In order to plot monthly costs predicted from the longitudinal models generated in the `R/models.R` script, we use the 
`plot_mean_monthly_costs()` function from the `R/functions.R` script to create the ggplot objects, and then save as `.svg` files.

```{r, eval=FALSE}

#Loop that plots the different outcomes longitudinally

for (outcome in c("grand_total_cost_month", "IN_CLM_month_total", "PS_REV_month_total", "IN_CLM_month_groupedHomeHealth",
                  "IN_CLM_month_groupedHospice","IN_CLM_month_groupedNonclaimauxiliary","IN_CLM_month_groupedDialysis",
                  "IN_CLM_month_groupedOutpatient","IN_CLM_month_groupedInpatient","IN_CLM_month_groupedSkilledNursingFacility")){
  
  
  plot_mean_monthly_costs(fit[[outcome]][["glmmTMB"]][["tweedie"]])
  ggsave(filename = paste0("figures/Longitudinal ", outcome," glmmTMB", " ", "tweedie.svg"),
         width=14,
         height=7
         )
  
  plot_mean_monthly_costs(fit[[outcome]][["glmmTMB"]][["linear"]])
  ggsave(filename = paste0("figures/Longitudinal ", outcome," glmmTMB", " ", "linear.svg"),
         width=14,
         height=7
  )
  
  
  plot_mean_monthly_costs(fit[[outcome]][["gee"]][["linear"]])
  ggsave(filename = paste0("figures/Longitudinal ", outcome," gee", " ", "linear.svg"),
         width=14,
         height=7
  )
  
  
  plot_mean_monthly_costs(fit[[outcome]][["glmmTMB"]][["log"]])
  ggsave(filename = paste0("figures/Longitudinal ", outcome," glmmTMB", " ", "log.svg"),
         width=14,
         height=7
  )
  
  
  plot_mean_monthly_costs(fit[[outcome]][["gee"]][["log"]])
  ggsave(filename = paste0("figures/Longitudinal ", outcome," gee", " ", "log.svg"),
         width=14,
         height=7
  )
  
  
}



```





## Other portions of the analysis

-   [**Setup**](setup.qmd): Defines global paths, data sources, cohort inclusion criteria, and analysis-wide constants.
-   [**Functions**](functions.qmd): Reusable helper functions for cohort construction, matching, costing, and modeling.
-   [**Create cohort**](create_cohort.qmd): Constructs the initial time-varying cohort of kidney transplant recipients, defining cohort entry, follow-up structure, and case/control eligibility prior to matching.
-   [**Execute matching**](execute_matching.qmd): Implements risk-set–based greedy matching without replacement to construct the analytic cohort.
-   [**Post-match processing**](postmatch_processing.qmd): Derives analytic variables, time-aligned cost windows, and follow-up structure after matching.
-   [**Modeling**](modeling.qmd): Fits prespecified cost and outcome models using the matched cohort.
-   [**Tables**](tables.qmd): Summary tables and regression outputs generated from the final models.
-   [**About**](about.qmd): methods, assumptions, and disclosures
