---
title: "Functions"
format: html
---

The code in this script generates tables for the manuscript.

::: {.Rcode title="Source code"}


The full R script is available at:

-   [`R/functions.R`](https://github.com/VagishHemmige/HOPE-CDC-analysis-2026/blob/master/R/functions.R)

This R script file is itself reliant on the following helper files:

-   [`R/setup.R`](https://github.com/VagishHemmige/HOPE-CDC-analysis-2026/blob/master/R/setup.R)
:::

## Drawing 60 minute isochrones around each transplant center

The following function uses the [`mapboxapi` package](https://walker-data.com/mapboxapi/) to calculate 60-minute travel isochrones around each
transplant center.

The function assumes the that you have an API token configured in R.  For details of how to do this, see the package documentation.

Its inputs are as follows:

- `SFObject`: This is an object of class `sf` that contains the location of each transplant center
- `year`: year as a character.  Must be a year in `yearlist` defined in `R/setup.R`.


```{r, eval=FALSE}



#Define a function that takes an SF data object and returns 60 minute isochrones
set_60_min_isochrone_at_7AM<-function(SFObject, year) {
  
  if (nrow(SFObject) == 0) {
    return(
      SFObject %>%
        st_transform(5070)
    )
  }else{
    tempDF<-SFObject%>%
      mutate(TimeZone=tz_lookup(., method="accurate"))%>%
      st_transform(4326)
    
    FinalDF <- vector("list", length(timezones))
    
    
    for(i in 1:length(timezones))
    {
      TZ<-converted_times[[year]][[i,1]]
      
      UTCTime<-format(converted_times[[year]][[i,3]], "%Y-%m-%dT%H:%M:%SZ")
      
      if (nrow(filter(tempDF,TimeZone==TZ))>0)
      {
        temp<-tempDF%>%
          filter(TimeZone==TZ)%>%
          mb_isochrone(time = 60, 
                       profile = "driving-traffic",
                       depart_at = UTCTime,
                       id_column = "OTCCode")%>%
          rename(OTCCode=id)
        FinalDF[[i]]<-temp
      }
    }
    
    FinalDF%>%
      bind_rows()%>%
      st_transform(5070)%>%
      return()
  }
}


```


## converting data frame of summary results into a `gt` format table

Inset LLM text here

```{r, eval=FALSE}

convert_resultsdf_to_table<-function(resultsdf, 
                                     year_list_fn=year_list,
                                     row_values,
                                     row_labels,
                                     table_title){
  
  
  resultsdf%>%
    
    #Reshape data set
    pivot_wider(
      names_from = Year,
      values_from = c(Numerator, Denominator, Percentage),
      names_vary = "slowest"
    )%>%
    
    #Use the table defined above to turn Characteristic into a factor variable
    mutate(Characteristic=factor(Characteristic,
                                 levels = row_values,
                                 labels = row_labels))%>%
    arrange(Characteristic)%>%
    
    #Convert to a gt format table
    gt()%>%
    
    #Groups columns by year with a spanner label
    {
      purrr::reduce(
        year_list_fn,
        ~ .x %>%
          tab_spanner(
            label = .y,
            columns = matches(paste0("_", .y, "$"))
          ),
        .init = .
      )
    } %>%
    
    #Sets coluns to round to 0 decimal places for numbers and 1 decimal place for percentages
    fmt_number(
      columns = matches(paste0("Numerator_|Denominator_")),
      decimals = 0,
      use_seps = TRUE
    ) %>%
    fmt_number(
      columns = matches("Percentage_"),
      decimals = 1
    )%>%tab_header(
      title = table_title
    )%>%
    
    #Removes "_year" from columns
    cols_label_with(~ "Numerator",   columns = matches("^Numerator_")) %>%
    cols_label_with(~ "Denominator", columns = matches("^Denominator_")) %>%
    cols_label_with(~ "Percentage",  columns = matches("^Percentage_"))%>%
    
    #Shades columns by year for easier reading
    
    {
      purrr::reduce(
        seq_along(year_list_fn),
        ~ .x %>%
          tab_style(
            style = cell_fill(
              color = colorRampPalette(c("lightgray","lightblue"))(
                length(year_list_fn)
              )[.y]            ),
            locations = cells_body(
              columns = matches(paste0("_", year_list_fn[.y], "$"))
            )
          ),
        .init = .
      )
    }%>%
    
    #Add separators between years
    {
      purrr::reduce(
        seq_len(length(year_list_fn) - 1),
        ~ .x %>%
          tab_style(
            style = cell_borders(
              sides = "right",
              weight = px(2),
              color = "black"
            ),
            locations = cells_body(
              columns = matches(paste0("Percentage_", year_list_fn[.y], "$"))
            )
          ) %>%
          tab_style(
            style = cell_borders(
              sides = "left",
              weight = px(2),
              color = "black"
            ),
            locations = cells_body(
              columns = matches(paste0("Numerator_", year_list_fn[.y + 1], "$"))
            )
          ),
        .init = .
      )
    }
  
  
}


}

```


## Creates paired plots of maps with percentages in the plot

LLM text

```{r, eval=FALSE}

make_paired_plot<-function(organ,
                           distance,
                           outcome,
                           buffer_list,
                           year1="2017",
                           year2="2022",
                           plottitle,
                           include_nonHIV=FALSE)
{
  
  # ----Error checking----
  
  # Validate include_nonHIV
  if (!is.logical(include_nonHIV) || length(include_nonHIV) != 1 || is.na(include_nonHIV)) {
    stop("include_nonHIV must be a single TRUE or FALSE value.")
  }
  
  
  #Access name of object passed to `buffer_list`
  buffer_name <- deparse(substitute(buffer_list))
  
  #Make sure name of outcome matches buffer_list
  if (!stringr::str_detect(
    stringr::str_to_lower(buffer_name),
    stringr::str_to_lower(outcome)
  )) {
    warning(glue::glue(
      "Buffer object ('{buffer_name}') may not match outcome '{outcome}'."
    ))
  }
  
  if (!year1 %in% year_list) {
    stop(glue::glue("year1 ({year1}) is not in year_list."))
  }
  
  if (!year2 %in% year_list) {
    stop(glue::glue("year2 ({year2}) is not in year_list."))
  }
  
  if (!outcome %in% c("active", "HIV", "HOPE")) {
    stop(glue::glue("outcome ({outcome}) is not in c('active', 'HIV', 'HOPE')."))
  }
  
  # ----Main function----
  
  #Keep only elements of Results_HIV_df necessary for plot
  summary_df_HIV<-Results_HIV_df[[organ]]%>%
    filter(str_detect(Characteristic, outcome))%>%
    filter(str_detect(Characteristic, distance))
  
  #parameters of plot assigned
  parameter_plot<-list()
  parameter_plot$y_title<-0.9
  parameter_plot$y_footnote<-0.1
  
  if (include_nonHIV)
  {
    parameter_plot$y_title<-0.97
    parameter_plot$y_footnote<-0.01
  }
  
  
  #Keep only elements of Results_nonHIV_df necessary for plot, if even needed in the first place
  if (include_nonHIV)
  {
    summary_df_nonHIV<-Results_nonHIV_df[[organ]]%>%
      filter(str_detect(Characteristic, outcome))%>%
      filter(str_detect(Characteristic, distance))
  }
  
  #Creates text for nonHIV (is NULL if )
  title_text_nonHIV<-list()
  
  if (include_nonHIV==TRUE) {
    title_text_nonHIV[[year1]]<-
      glue("<br>",
           "<span style='font-size:30pt; font-weight:normal;'>HIV negative:</span> ",
           "<span style='font-size:30pt; font-weight:bold;'>",
           "{summary_df_nonHIV$Percentage[summary_df_nonHIV$Year == as.numeric(year1)]}%",
           "</span>",
           "<br>",
           "{comma(summary_df_nonHIV$Numerator[summary_df_nonHIV$Year == as.numeric(year1)])}",
           " / ",
           "{comma(summary_df_nonHIV$Denominator[summary_df_nonHIV$Year == as.numeric(year1)])}")
    
    title_text_nonHIV[[year2]]<-
      glue("<br>",
           "<span style='font-size:30pt; font-weight:normal;'>HIV negative:</span> ",
           "<span style='font-size:30pt; font-weight:bold;'>",
           "{summary_df_nonHIV$Percentage[summary_df_nonHIV$Year == as.numeric(year2)]}%",
           "</span>",
           "<br>",
           "{comma(summary_df_nonHIV$Numerator[summary_df_nonHIV$Year == as.numeric(year2)])}",
           " / ",
           "{comma(summary_df_nonHIV$Denominator[summary_df_nonHIV$Year == as.numeric(year2)])}")
  }
  else {
    title_text_nonHIV[[year1]]<-""
    title_text_nonHIV[[year2]]<-""
  }
  
  
  
  #Left graph
  Plot1<-ggplot() +
    geom_sf(data = county_dots_transformed[[year1]], size = 0.3, color = "navy", alpha = 0.5) +
    theme_void() +
    theme(
      axis.title.y = element_text(size = 36, angle = 90, vjust = 0.5),
      plot.title = element_textbox(size = 16, hjust = 0.5,
                                   halign = 0.5)
    )+
    labs(y=year1,
         title = glue("<span style='font-size:30pt; font-weight:normal;'>PLWH:</span> ",
                      "<span style='font-size:30pt; font-weight:bold;'>",
                      "{summary_df_HIV$Percentage[summary_df_HIV$Year == as.numeric(year1)]}%",
                      "</span>",
                      "<br>",
                      "{comma(summary_df_HIV$Numerator[summary_df_HIV$Year == as.numeric(year1)])}",
                      " / ",
                      "{comma(summary_df_HIV$Denominator[summary_df_HIV$Year == as.numeric(year1)])}",
                      title_text_nonHIV[[year1]]
         ))+
    geom_sf(data=buffer_list[[organ]][[distance]][[year1]], color="red", fill=NA)+
    geom_sf(data=states_sf_transformed, fill=NA)
  
  
  #Right plot
  Plot2<-ggplot() +
    geom_sf(data = county_dots_transformed[[year2]], size = 0.3, color = "navy", alpha = 0.5) +
    theme_void() +
    theme(
      axis.title.y = element_text(size = 36, angle = 90, vjust = 0.5),
      plot.title = element_textbox(size = 16, hjust = 0.5,
                                   halign = 0.5)
    )+
    labs(y=year2,
         title = glue("<span style='font-size:30pt; font-weight:normal;'>PLWH:</span> ",
                      "<span style='font-size:30pt; font-weight:bold;'>",
                      "{summary_df_HIV$Percentage[summary_df_HIV$Year == as.numeric(year2)]}%",
                      "</span>",
                      "<br>",
                      "{comma(summary_df_HIV$Numerator[summary_df_HIV$Year == as.numeric(year2)])}",
                      " / ",
                      "{comma(summary_df_HIV$Denominator[summary_df_HIV$Year == as.numeric(year2)])}",
                      title_text_nonHIV[[year2]]
         ))+
    geom_sf(data=buffer_list[[organ]][[distance]][[year2]], color="red", fill=NA)+
    geom_sf(data=states_sf_transformed, fill=NA)
  
  #Combine plots
  combined<-cowplot::plot_grid(Plot1, Plot2,
                               ncol = 2, align = "hv")
  
  final_plot <- ggdraw() +
    draw_plot(combined) +
    draw_label(
      plottitle,
      x = 0.5, y = parameter_plot$y_title,   # top center
      hjust = 0.5, vjust = 1,   # anchor to the top edge
      fontface = 'bold',
      size = 25
    )+
    # 
    draw_label(
      "Each blue dot represents 100 people with HIV\nRed outlines represent the catchment area for each transplant center",
      x = 0.02, y = parameter_plot$y_footnote,          # left margin position
      hjust = 0, vjust = 0,        # align text to the left edge
      size = 14,
      fontface = "italic"
    )
  
  return(final_plot)
  
}

```


## Plot the number of people with HIV in the catchment area of each transplant center

This function uses the `ggbeeswarm` package to make GraphPad Prism-type plots comparing the number of people living with HIV in the catchment area of each transplant center in the years specified.

```{r, eval=FALSE}


plot_center_HIV_catchment<-function(organ,
                                    distance,
                                    buffer_list,
                                    year1="2017",
                                    year2="2022"){
  
  
  
  # ----Error checking----
  
  #Access name of object passed to `buffer_list`
  buffer_name <- deparse(substitute(buffer_list))
  
  if (!year1 %in% year_list) {
    stop(glue::glue("year1 ({year1}) is not in year_list."))
  }
  
  if (!year2 %in% year_list) {
    stop(glue::glue("year2 ({year2}) is not in year_list."))
  }
  
  
  
  temp_DF<-list()
  
  temp_DF[[year1]]<-buffer_list%>%
    pluck(year1)%>%
    st_drop_geometry()%>%
    group_by(OTCCode)%>%
    summarize(total_cases=round(sum(tract_cases), 0))%>%
    mutate(year=as.numeric(year1))%>%
    mutate(Center_classification=case_when(
      OTCCode %in% HOPE_center_volumes[[organ]][[year1]]$REC_CTR_CD~"HOPE center",
      OTCCode %in% HIV__center_volumes[[organ]][[year1]]$REC_CTR_CD~"≥1 HIV R+ transplants",
      OTCCode %in% Active_tx_centers[[organ]][[year1]]~"No HIV R+ transplants"
    ))%>%
    filter(!is.na(OTCCode))
  
  
  temp_DF[[year2]]<-buffer_list%>%
    pluck(year2)%>%
    st_drop_geometry()%>%
    group_by(OTCCode)%>%
    summarize(total_cases=round(sum(tract_cases), 0))%>%
    mutate(year=as.numeric(year2))%>%
    mutate(Center_classification=case_when(
      OTCCode %in% HOPE_center_volumes[[organ]][[year2]]$REC_CTR_CD~"HOPE center",
      OTCCode %in% HIV__center_volumes[[organ]][[year2]]$REC_CTR_CD~"≥1 HIV R+ transplants",
      OTCCode %in% Active_tx_centers[[organ]][[year2]]~"No HIV R+ transplants"
    ))%>%
    filter(!is.na(OTCCode))
  
  combined_DF<-bind_rows(temp_DF)%>%
    mutate(Center_classification=factor(Center_classification,
                                        levels=c("HOPE center","≥1 HIV R+ transplants", "No HIV R+ transplants"),
                                        labels=c("HOPE center","≥1 HIV R+ transplants", "No HIV R+ transplants")
    ))%>%
    mutate(Center_classification_color=case_when(
      OTCCode %in% HOPE_center_volumes[[organ]][[year1]]$REC_CTR_CD~"HOPE center",
      OTCCode %in% HIV__center_volumes[[organ]][[year1]]$REC_CTR_CD~"≥1 HIV R+ transplants",
      OTCCode %in% Active_tx_centers[[organ]][[year1]]~"No HIV R+ transplants",
      TRUE~glue("No transplants in {year1}")
    ))
  
  combined_DF %>% ggplot()+
    geom_violin(mapping = aes(x=Center_classification, y=total_cases), color = NA, alpha=0.5,fill = "grey", 
                draw_quantiles = 1)+
    stat_summary(mapping = aes(x=Center_classification, y=total_cases),fun = median, 
                 fun.min = median, fun.max = median, 
                 geom = "crossbar", width = 0.9, fatten=1.5)+
    geom_quasirandom(mapping = aes(x=Center_classification, y=total_cases, color=Center_classification_color), 
                     cex=2, alpha=0.7)+
    stat_compare_means(method = "kruskal.test",
                       mapping = aes(x = Center_classification, y = total_cases)) +
    scale_y_continuous(labels = scales::comma)+
    guides(color = guide_legend(title.position = "top", nrow = 1))+
    theme_classic()+
    theme(legend.position = "bottom",
          legend.title = element_text(hjust = 0.5),  # centers title
          legend.box = "horizontal")+
    theme(axis.text.x = element_text(
      angle = 45,
      hjust = 1,
      vjust = 1))+
    facet_grid(.~year)+
    labs(x = "Center HIV R+ volume", y="PLWH in catchment area", color=glue("{year1} status"))+
    scale_color_discrete(breaks = c("HOPE center",
                                    "≥1 HIV R+ transplants", 
                                    "No HIV R+ transplants",
                                    glue("No transplants in {year1}"))
    )
  
  
  
}
```


## Other portions of the analysis


-   [**Setup**](setup.qmd): Defines global paths, data sources, cohort inclusion criteria, and analysis-wide constants.
-   [**Tables**](tables.qmd): Summary tables and regression outputs generated from the final models.
-   [**Figures**](figures.qmd):Visualizations of costs, risks, and model-based estimates.
-   [**About**](about.qmd): methods, assumptions, and disclosures
