---
title: "Results: kidney"
format: html
---

## Table 1 and Table 2: Summary results in PLWH and without HIV


```{r, include=FALSE}

#This code prepares the .qmd environent for analysis

organ<-"heart"

Results_HIV_gt    <- readRDS("../cache/Results_HIV_gt.rds")
Results_nonHIV_gt <- readRDS("../cache/Results_nonHIV_gt.rds")

source("../R/setup.R")


```



### Table 1: People Living With HIV (PLWH)

This table summarizes the number and proportion of people living with HIV who reside within defined catchment areas of:

- Active transplant centers  
- HIV R+ transplant centers  
- HOPE Act D+ centers  

Catchment areas are defined by:

- 50-mile radius  
- 60-minute driving time  
- 100-mile radius  
- 200-mile radius  

Results are shown for 2017 and 2022.

---

```{r, echo=FALSE}

Results_HIV_gt[[organ]]

```

### Table 2: Individuals Without HIV

This table presents the same catchment analysis for individuals without HIV.  

It allows direct comparison of geographic access between:

- PLWH  
- Individuals without HIV  

---

```{r, echo=FALSE}


Results_nonHIV_gt[[organ]]

```



## Figures

Figures are grouped into three sections.

---

### Compare PLWH and non-HIV

These figures compare geographic access between PLWH and individuals without HIV within each catchment definition.

Each panel corresponds to a distance-based definition:

- 50 miles  
- 60 minutes  
- 100 miles  
- 200 miles  

These figures highlight disparities in access by HIV status.

---

```{r, echo=FALSE}


list.files(
  glue::glue("../figures/{organ}"),
  pattern = "Combined.*\\.png$",
  full.names = TRUE
)%>%
  str_subset("Active")%>%
  {
    .[order(as.numeric(stringr::str_extract(., "\\d+(?=(mile|min))")))]
  } %>%
  knitr::include_graphics()

```

---

### Compare PLWH Between Years

These figures evaluate changes in geographic access among PLWH between 2017 and 2022.

For each catchment definition, results are shown for:

- Active centers  
- HIV R+ centers  
- HOPE Act D+ centers  

This section highlights temporal changes in transplant center availability.

---


```{r, echo=FALSE}


list.files(
  glue::glue("../figures/{organ}"),
  pattern = "Combined.*\\.png$",
  full.names = TRUE
)%>%
  str_subset("Active", negate=TRUE)%>%
  {
    .[order(as.numeric(stringr::str_extract(., "\\d+(?=(mile|min))")))]
  } %>%
  knitr::include_graphics()

```

---

### Beeswarm Plots

Beeswarm plots display the distribution of catchment-level population proportions across centers.

Each point represents a transplant center, with the distribution shown for each distance definition.

These plots illustrate variability in geographic access across centers rather than only aggregate proportions.

```{r, echo=FALSE}


list.files(
  glue::glue("../figures/{organ}"),
  pattern = "Beeswarm.*\\.png$",
  full.names = TRUE
)%>%
  {
    .[order(as.numeric(stringr::str_extract(., "\\d+(?=(mile|min))")))]
  } %>%
  knitr::include_graphics()

```

